---
title: "penguinPower"
author: "Valarie"
date: '2022-07-04'
output:
  html_document: default
  pdf_document: default
---

# Outline

The intention here is to provide a comparative view between **3** different species of penguins:

* Adelie
* Gentoo
* Chinstrap

I will begin by setting our environment in R, so that the code chunks can successfully run. 

Then, using our data and respective code chunks, I will provide a plot chart for us to examine trends between the respective penguins and their _flipper length_ to _body mass_ ratio. This information will allow us to determine which species are, on average, smallest or largest.

Finally, I will offer a conclusion given the data we've analyzed in this report.

## Setting the environment

Here, I will be loading in some R packages to assist with the visualization, as well as our dataset itself.

```{r loading packages, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(palmerpenguins)
```

Now that our packages are installed, let's move on to the chart.

### Creating a plot chart

Here, I will insert a code chunk that creates a solid visual representation of our 3 penguin species and their respective sizes. 
This chart will utilize both a scatter plot representation and a smooth line graph to help us assess trends.

```{r generate graph, message=FALSE, warning=FALSE}
ggplot(data=penguins) +
  geom_smooth(mapping=aes(x=flipper_length_mm, y=body_mass_g), color="black") +
  geom_point(mapping=aes(x=flipper_length_mm, y=body_mass_g, color=species)) +
  labs(title="Palmer's Penguins: Flipper Length vs. Body Mass", 
       subtitle="A comparative view between 3 species",
       caption="Data from Palmer Penguins dataset")
```

Here, we can see quite clearly that, in terms of _both_ flipper length and body mass, the Gentoo species is, on average, the largest.
Chipstrap and Adelie are a bit closer of a matchup, but it would appear that the Adelie species is marginally larger, on average.

#### Conclusion

It is safe to conclude that the Gentoo species is the all around largest species between the 3 populations. I would feel comfortable in claiming that Chinstrap are the smallest across the board, but it is absolutely worth noting how minimal that margin is from the Adelie species.